Managing variability in software systems

### 2021-05-02 Paper.tex
decision modeling

Decision models can be represented in different ways, broadly used approaches are flat lists or tables of decisions which also includes their dependencies. In decision modeling hierarchy is a secondary aspect, because there are no uniform rules to describe hierarchy. \cite{b6}\\

@@

@@

For further information we refer the reader to the references, especially \cite{b2}, which gives a more in depth overview, and \cite{b6}, which compares different approaches to variability modelling.
\newpage

\begin{thebibliography}{00}
\bibitem{b1} S. Apel, D. Batory, C. Kästner and G. Saake, "Software Product Lines. In: Feature-Oriented Software Product Lines", Berlin Heidelberg, Springer (2013)
\bibitem{b2} J. van Gurp, J. Bosch and M. Svahnberg, "On the notion of variability in software product lines," Proceedings Working IEEE/IFIP Conference on Software Architecture", Amsterdam, Netherlands, 2001, pp. 45-54
\bibitem{b3} F. Bachmann, L. Bass, "Managing Variability in Software Architectures", Carnegie Mellon University (2001)
\bibitem{b4} M. Galster, D. Weyns, D. Tofan, M. Bartosz, P. Avgeriou, "Variability in Software Systems—A Systematic Literature Review", IEEE Transactions on Software Engineering (2014)
\bibitem{b5} T. Berger, J.-P. Steghöfer, T. Ziadi, J. Robin, and J. Martinez, "The state of adoption and the challenges of systematic variability management in industry", Empirical Software Engineering, vol. 25, pp. 1755-1797, 2020
\bibitem{b6}
K. Czarnecki, P. Grünbacher, R. Rabiser, K. Schmid., and A.
Wasowski, “Cool features and tough decisions: a comparison of
variability modeling approaches,” Proc. of the \textit{Sixth international
workshop on variability modeling of software-intensive systems
(VaMoS '12)}, ACM, Jan 2012, pp. 173-182.
\bibitem{b7}
Software Productivity Consortium Services Corporation,
Technical Report SPC-92019-CMC. \textit{Reuse-Driven Software
Processes Guidebook}, Version 02.00.03, 1993.
\bibitem{b8} K. C. Kang, S. Cohen, J. Hess, W. Novak, and S. Peterson,	 "Feature oriented domain analysis (FODA) feasibility study", Technical Report CMU/SEI-90-TR-21, Software Engineering Institute, Carnegie Mellon University, Pittburgh, PA, USA, 1990
\bibitem{b9} A. Metzger, and K. Pohl, "Software product line engineering and variability management: Achievements and challenges," \textit{Proc. of the Future of Software Engineering (FOSE 2014)}, ACM, 2014, pp. 70-84, doi:10.1145/2593882.2593888.
\bibitem{b10} M. Riebisch, K. Böllert, D. Streitferdt, and I. Philippow, "Extending Feature Diagrams with UML Multiplicities", \textit{6th World Conference on Integrated Design \& Process Technology (IDPT2002)}, vol. 23, 2002, pp. 1-7.
\end{thebibliography}

>>>>

Decision models can be represented in different ways, broadly used approaches are flat lists or tables of decisions which also includes their dependencies. In decision modeling hierarchy is a secondary aspect, because there are no uniform rules to describe hierarchy.\cite{b6}\\

@@

As mentioned in \cite{b11} the  different approaches for decision modeling have some common core model structure. There is an entity which is called decision, which represents a set of all possible choices which are available at a certain point in time.\\

These decision have attributes which can vary from one modeling approach to another. But commonly at least a decision name or identifier are used, while the other atrributes are describing the decision in such a way that the user can understand it. Furthermore the decision has a data type which defines the set of valid choices for decision.\cite{b11}\\

With all of the different approaches it is possible to create dependencies between decisions. There are value contraints for defining dependencies between values, which are supported by most approaches. But here it has to be mentioned that all the approaches use different concepts for value contraints and generally speaking for the management of the dependencies.\cite{b11}\\

Some of them are using contraints, similar to the approach used in feature modeling. Another approach for modeling dependencies is with rules or decisions, which are are checked when a decision is taken.\cite{b11} \\

The second approach could be more intuitive for most people with a software engineering background because of the similarity how a computer program with conditional statements works.

@@

For further information we refer the reader to the references, especially \cite{b2}, which gives a more in depth overview, and \cite{b6}, which compares different approaches to variability modelling.

\begin{thebibliography}{00}
\bibitem{b1} S. Apel, D. Batory, C. Kästner and G. Saake, "Software Product Lines. In: Feature-Oriented Software Product Lines", Berlin Heidelberg, Springer (2013)
\bibitem{b2} J. van Gurp, J. Bosch and M. Svahnberg, "On the notion of variability in software product lines," Proceedings Working IEEE/IFIP Conference on Software Architecture", Amsterdam, Netherlands, 2001
\bibitem{b3} F. Bachmann, L. Bass, "Managing Variability in Software Architectures", Carnegie Mellon University (2001)
\bibitem{b4} M. Galster, D. Weyns, D. Tofan, M. Bartosz, P. Avgeriou, "Variability in Software Systems—A Systematic Literature Review", IEEE Transactions on Software Engineering (2014)
\bibitem{b5} T. Berger, J.-P. Steghöfer, T. Ziadi, J. Robin, and J. Martinez, "The state of adoption and the challenges of systematic variability management in industry", Empirical Software Engineering, vol. 25, 2020
\bibitem{b6}
K. Czarnecki, P. Grünbacher, R. Rabiser, K. Schmid., and A.
Wasowski, “Cool features and tough decisions: a comparison of
variability modeling approaches”, Proc. of the \textit{Sixth international
workshop on variability modeling of software-intensive systems
(VaMoS '12)}, ACM, Jan 2012
\bibitem{b7}
Software Productivity Consortium Services Corporation,
Technical Report SPC-92019-CMC. \textit{Reuse-Driven Software
Processes Guidebook}, Version 02.00.03, 1993.
\bibitem{b8} K. C. Kang, S. Cohen, J. Hess, W. Novak, and S. Peterson,	 "Feature oriented domain analysis (FODA) feasibility study", Technical Report CMU/SEI-90-TR-21, Software Engineering Institute, Carnegie Mellon University, Pittburgh, PA, USA, 1990
\bibitem{b9} A. Metzger, and K. Pohl, "Software product line engineering and variability management: Achievements and challenges," \textit{Proc. of the Future of Software Engineering (FOSE 2014)}, ACM, 2014, pp. 70-84, doi:10.1145/2593882.2593888.
\bibitem{b10} M. Riebisch, K. Böllert, D. Streitferdt, and I. Philippow, "Extending Feature Diagrams with UML Multiplicities", \textit{6th World Conference on Integrated Design \& Process Technology (IDPT2002)}, vol. 23, 2002
\bibitem{b11} K. Schmid, R. Rabiser, P. Grünbacher, "A comparison of decision modeling approaches in product lines", Proc. of the \textit{Fifth international
workshop on variability modeling of software-intensive systems
(VaMoS '11)}, ACM, Jan 2011
\end{thebibliography}
